---
title: "exercise-12"
output: html_document
date: "2024-04-13"
---

```{r, warning = FALSE}
library(tidyverse)
library(dplyr)
library(mosaic)
library(readr)
library(skimr)
library(broom)
library(car)
library(manipulate)
library(ggplot2)
library(infer)
library(modelr)
```

```{r}
#Read in Mammal Life Histories dataset
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f, col_names = TRUE)
```

## Exploratory data analysis

In the dataset, d:

* the following variables (total = 10) are numeric:

```{r}
skim <- skim(d)

skim_num <- skim |> filter(skim_type == "numeric")
print(skim_num$skim_variable) #Which variables are numeric
print(length(skim_num$skim_variable)) #How many variables are numeric
```

...and the following variables (total = 4) are categorical:

```{r}
skim_char <- skim |> filter(skim_type == "character")
print(skim_char$skim_variable) #Which are character variables
print(length(skim_char$skim_variable)) #How many are character variables
```

## Step 1

Replace all values of -999 (the authors’ code for missing data) with NA.

```{r}
#Replace -999s with NA
d <- d |> mutate(
  `mass(g)` = na_if(`mass(g)`, -999),
  `gestation(mo)` = na_if(`gestation(mo)`, -999),
  `newborn(g)` = na_if(`newborn(g)`, -999),
  `weaning(mo)`= na_if(`weaning(mo)`, -999),
  `wean mass(g)` = na_if(`wean mass(g)`, -999),
  `AFR(mo)` = na_if(`AFR(mo)`, -999),
  `max. life(mo)` = na_if(`max. life(mo)`, -999),
  `litter size` = na_if(`litter size`, -999),
  `litters/year` = na_if(`litters/year`, -999))
```

## Step 2

Drop the variables litter size and refs.

```{r}
d <- d |>  select(-`litter size`, -`refs`)
```

## Step 3

Log transform all of the other numeric variables.

```{r}
d <- d |> mutate(across(where(is.numeric), log))
```

## Step 4

Regress the (now log transformed) age [gestation(mo), weaning(mo), AFR(mo) (age at first reproduction), and max. life(mo) (maximum lifespan)] and mass [newborn(g) and wean mass(g)] variables on (now log transformed) overall body mass(g) and add the residuals to the dataframe as new variables (relGest, relWean, relAFR, relLife, relNewbornMass, and relWeaningMass).

Then, add the residuals to the dataframe as new variables (relGest, relWean, relAFR, relLife, relNewbornMass, and relWeaningMass).

```{r}
#Regressing gestation(mo) on overall body mass(g)

m1 <- lm(`gestation(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m1))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relGest)

d$relGest <- residuals(m1)
```

```{r}
#Regressing weaning(mo) on overall body mass(g)

m2 <- lm(`weaning(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m2))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relWean)

d$relWean <- residuals(m2)
```

```{r}
#Regressing AFR(mo) on overall body mass(g)

m3 <- lm(`AFR(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m3))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relAFR)

d$relAFR <- residuals(m3)
```

```{r}
#Regressing max.life(mo) on overall body mass(g)

m4 <- lm(`max. life(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m4))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relLife)

d$relLife <- residuals(m4)
```

```{r}
#Regressing newborn(g) on overall body mass(g)

m5 <- lm(`newborn(g)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m5))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relNewbornMass)

d$relNewbornMass <- residuals(m5)
```

```{r}
#Regressing wean mass(g) on overall body mass(g)

m6 <- lm(`wean mass(g)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m6))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relWeaningMass)

d$relWeaningMass <- residuals(m6)
```

## Step

Plot residuals of max lifespan (relLife) in relation to Order. 

```{r}
pm1 <- ggplot(data = d, aes(x = order, y = relLife)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Residuals of Max Lifespan in Relation to Order ") +
  ylab("Residuals") + 
  xlab("Order") 

(pm1)
```

Which mammalian orders have the highest residual lifespan?

Plot residuals of max lifespan (relLife) in relation to Order. 

```{r}

```

Which mammalian orders have the highest residual lifespan?

Plot residuals of newborn mass (relNewbornMass) in relation to Order. 

```{r}

```

Which mammalian orders have the have highest residual newborn mass?

Plot residuals of weaning mass (relWeaningMass) in relation to Order.

```{r}

```

Which mammalian orders have the have highest relative weaning mass?

