---
title: "exercise-12"
output: html_document
date: "2024-04-13"
---

```{r, warning = FALSE}
library(tidyverse)
library(dplyr)
library(mosaic)
library(readr)
library(skimr)
library(broom)
library(car)
library(manipulate)
library(ggplot2)
library(infer)
library(modelr)
library(MASS)
library(AICcmodavg)
```

```{r}
#Read in Mammal Life Histories dataset
f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/Mammal_lifehistories_v2.txt"
d <- read_tsv(f, col_names = TRUE)
```

## Exploratory data analysis

In the dataset, d:

* the following variables (total = 10) are numeric:

```{r}
skim <- skim(d)

skim_num <- skim |> filter(skim_type == "numeric")
print(skim_num$skim_variable) #Which variables are numeric
print(length(skim_num$skim_variable)) #How many variables are numeric
```

...and the following variables (total = 4) are categorical:

```{r}
skim_char <- skim |> filter(skim_type == "character")
print(skim_char$skim_variable) #Which are character variables
print(length(skim_char$skim_variable)) #How many are character variables
```

## Step 1

Replace all values of -999 (the authors’ code for missing data) with NA.

```{r}
#Replace -999s with NA
d <- d |> mutate(
  `mass(g)` = na_if(`mass(g)`, -999),
  `gestation(mo)` = na_if(`gestation(mo)`, -999),
  `newborn(g)` = na_if(`newborn(g)`, -999),
  `weaning(mo)`= na_if(`weaning(mo)`, -999),
  `wean mass(g)` = na_if(`wean mass(g)`, -999),
  `AFR(mo)` = na_if(`AFR(mo)`, -999),
  `max. life(mo)` = na_if(`max. life(mo)`, -999),
  `litter size` = na_if(`litter size`, -999),
  `litters/year` = na_if(`litters/year`, -999))
```

## Step 2

Drop the variables litter size and refs.

```{r}
d <- d |>  dplyr::select(-`litter size`, -`refs`)
```

## Step 3

Log transform all of the other numeric variables.

```{r}
d <- d |> mutate(across(where(is.numeric), log))
```

## Step 4

Regress the (now log transformed) age [gestation(mo), weaning(mo), AFR(mo) (age at first reproduction), and max. life(mo) (maximum lifespan)] and mass [newborn(g) and wean mass(g)] variables on (now log transformed) overall body mass(g) and add the residuals to the dataframe as new variables (relGest, relWean, relAFR, relLife, relNewbornMass, and relWeaningMass).

Then, add the residuals to the dataframe as new variables (relGest, relWean, relAFR, relLife, relNewbornMass, and relWeaningMass).

```{r}
#Regressing gestation(mo) on overall body mass(g)

m1 <- lm(`gestation(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m1))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relGest)

d$relGest <- residuals(m1)
```

```{r}
#Regressing weaning(mo) on overall body mass(g)

m2 <- lm(`weaning(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m2))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relWean)

d$relWean <- residuals(m2)
```

```{r}
#Regressing AFR(mo) on overall body mass(g)

m3 <- lm(`AFR(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m3))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relAFR)

d$relAFR <- residuals(m3)
```

```{r}
#Regressing max.life(mo) on overall body mass(g)

m4 <- lm(`max. life(mo)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m4))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relLife)

d$relLife <- residuals(m4)
```

```{r}
#Regressing newborn(g) on overall body mass(g)

m5 <- lm(`newborn(g)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m5))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relNewbornMass)

d$relNewbornMass <- residuals(m5)
```

```{r}
#Regressing wean mass(g) on overall body mass(g)

m6 <- lm(`wean mass(g)` ~ `mass(g)`, data = d, na.action = na.exclude)
(summary(m6))
```

```{r}
#Adding residuals to dataframe (d) as a new variable (relWeaningMass)

d$relWeaningMass <- residuals(m6)
```

## Step

Plot residuals of max lifespan (relLife) in relation to Order. 

```{r}
p1 <- ggplot(data = d, aes(x = order, y = relLife)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Residuals of Max Lifespan in Relation to Order ") +
  ylab("Residuals") + 
  xlab("Order") 
(p1)
```

Order Primates has the highest residual lifespan


Plot residuals of newborn mass (relNewbornMass) in relation to Order. 

```{r}
p2 <- ggplot(data = d, aes(x = order, y = relNewbornMass)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Residuals of Newborn Mass in Relation to Order ") +
  ylab("Residuals") + 
  xlab("Order") 
(p2)
```

Order Macroscelidea has the highest residual newborn mass.


Plot residuals of weaning mass (relWeaningMass) in relation to Order.

```{r}
p3 <- ggplot(data = d, aes(x = order, y = relWeaningMass)) + 
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90)) +
  ggtitle("Residuals of Weaning Mass in Relation to Order ") +
  ylab("Residuals") + 
  xlab("Order") 
(p3)
```

Order Perissodactyla has the highest residual weaning mass.

## Step 6

Run models and a model selection process to evaluate what variables best predict max. life(mo) from the set of the following predictors: gestation(mo), newborn(g), weaning(mo), wean mass(g), litters/year, and overall body mass(g).

```{r}
d_new <- d |> drop_na(`max. life(mo)`,`AFR(mo)`,`gestation(mo)`,`newborn(g)`,`weaning(mo)`,`wean mass(g)`,`litters/year`,`mass(g)`)

m_full1 <- lm(data = d_new, `max. life(mo)` ~ `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `wean mass(g)` + `litters/year` + `mass(g)`)

s1 <- stepAIC(m_full1, scope = .~., direction = "both", trace = TRUE)
(s1)
```

Based on AIC values calculated in the above step-wise model selection using {MASS}, the variables that best predict max. life(mo) are gestation(mo), weaning(mo), litters/year, and mass(g).

```{r}
# I am not sure if there is a more efficient way to do this rather than doing a model selection process by hand?

#Backward selection

ma <- lm(`max. life(mo)` ~ `gestation(mo)` + `newborn(g)` + `weaning(mo)` + 
    `wean mass(g)` + `litters/year` + `mass(g)`, data = d_new)

drop1(ma, test = "F")

mb <- update(ma, . ~ . - `weaning(mo)`)

drop1(mb, test = "F")

mc <- update(mb, . ~ . - `wean mass(g)`)

drop1(mc, test = "F")

md <- update(mc, . ~ . - `newborn(g)`)

drop1(md, test = "F")

me <- update(md, . ~ . - `mass(g)`)

drop1(me, test = "F")

mf <- update(me, . ~ . - `litters/year`)

drop1(mf, test = "F")

mg <- update(mf, . ~ . - `gestation(mo)`)

(aictab(list(ma, mb, mc,md,me,mf,mg), c("ma", "mb", "mc","md","me","mf","mg")))
```

The best model overall based on AICc is:

max. life(mo) ~ gestation(mo) + newborn(g) + weaning(mo) + wean mass(g) + litters/year + mass(g)

4 models have a delta AICc of 4 or less. Gestation(mo), litters/year, and mass(g) appear in all "top" models.

---

Run models and a model selection process to evaluate what variables best predict AFR(mo) from the set of the following predictors: gestation(mo), newborn(g), weaning(mo), wean mass(g), litters/year, and overall body mass(g).

```{r}
m_full2 <- lm(data = d_new, `AFR(mo)` ~ `gestation(mo)` + `newborn(g)` + `weaning(mo)` + `wean mass(g)` + `litters/year` + `mass(g)`)

s2 <- stepAIC(m_full2, scope = .~., direction = "both", trace = TRUE)
(s2)
```

Based on AIC values calculated in the above step-wise model selection using {MASS}, the variables that best predict AFR(mo) are gestation(mo), weaning(mo), litters/year, and mass(g).

```{r}
#Backward selection

ma2 <- lm(`AFR(mo)` ~ `gestation(mo)` + `newborn(g)` + `weaning(mo)` + 
    `wean mass(g)` + `litters/year` + `mass(g)`, data = d_new)

drop1(ma2, test = "F")

mb2 <- update(ma2, . ~ . - `newborn(g)`)

drop1(mb2, test = "F")

mc2 <- update(mb2, . ~ . - `wean mass(g)`)

drop1(mc2, test = "F")

md2 <- update(mc2, . ~ . - `mass(g)`)

drop1(md2, test = "F")

me2 <- update(md2, . ~ . - `weaning(mo)`)

drop1(me2, test = "F")

mf2 <- update(me2, . ~ . - `gestation(mo)`)

drop1(mf2, test = "F")

mg2 <- update(mf2, . ~ . - `litters/year`)

(aictab(list(ma2, mb2, mc2,md2,me2,mf2,mg2), c("ma2", "mb2", "mc2","md2","me2","mf2","mg2")))
```

The best model overall based on AICc is:

AFR(mo) ~ gestation(mo) + weaning(mo) + litters/year + mass(g)

4 models have a delta AICc of 4 or less. Gestation(mo) and litters/year appear in all "top" models.

---

Calculate and plot the model-averaged coefficients and their CIs across this set of top models.


